---
title: "Final Project"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r required packages}
library(tidyverse) #cleaning data ect.
library(lme4) #fitting mixed models
library(glmmTMB) #not all models fit with lme4, some use this package
library(DHARMa) #Used for model analysis with glmmTMB models. 
library(emmeans) #Used to look at model contrasts 
library(car) #Stats 
library(effects) #to plot some model effects 
library(dotwhisker) #creating dot whisker plots 
library(broom) #some tidying functions used for code from course website. 
library(corrplot) #looking at corralations 
```

```{r required functions}
PC_plot <- function(dat = gene1_combined, x = wings_combined$PC1, y = wings_combined$PC1) {
  ggplot(dat, aes(x = x, y = y, col = background, shape = sex)) +
    geom_point(alpha = 0.4) 
}

size_cor <- function(dat = gene1_combined, PC = wings_combined$PC1) {
  ggplot(dat, aes(x = WL, y = PC, col = background)) +
    geom_point(alpha = 0.4) +
    geom_smooth(method = lm)
}

bootie <- function(dat=gene1only, genename = "Egfr", vein = gene1only$L2s) {
  new_dat <- dat[sample(nrow(dat), nrow(dat), replace = T),  ]
  mod_new <- lm(vein ~ background*sex, 
                data = new_dat,
                subset = gene1only$gene1 == genename)
  return(coef(mod_new))
}

bootie_mm <- function(dat=gene1only_long, genename = "Egfr") {
  new_dat <- dat[sample(nrow(dat), nrow(dat), replace = T),  ]
  mod_new <- lm(scalelength ~ vein:background*sex, 
                data = new_dat,
                subset = gene1only$gene1 == genename)
  return(coef(mod_new))
}

bootie_mlm <-  function(dat=gene1only, genename = "Egfr") {
  new_dat <- gene1only[sample(nrow(gene1only), nrow(gene1only), replace = T),  ]
  mod_new <- lm(as.matrix(new_dat[,13:16]) ~ background*sex, 
                data = new_dat,
                subset = gene1only$gene1 == genename)
  return(coef(mod_new))
}

conf_bootie_f <- function(dat = gene1only, genename = "Egfr", vein = gene1only$L2s, fun = bootie){
  coef_out <- replicate(n = 100, fun(dat = dat, genename = genename, vein = vein))
  ratio_coef_f <- coef_out[2,]/coef_out[1,]
  CI_f <- quantile(ratio_coef_f, probs = c(0.05, 0.5, 0.95))
  print(CI_f)
}
conf_bootie_m <- function(dat = gene1only, genename = "Egfr", vein = gene1only$L2s, fun = bootie){
  coef_out <- replicate(n = 100, fun(dat = dat, genename = genename, vein = vein))
  ratio_coef_m <- coef_out[4,]/coef_out[3,]
  quantile(ratio_coef_m, probs = c(0.05, 0.5, 0.95))
}


conf_mm_f <- function(dat = gene1only, genename = "Egfr", fun = bootie){
  coef_out <- replicate(n = 100, fun(dat = dat, genename = genename))
  ratio_coef_f <- coef_out[3:5]/coef_out[7:9]
  CI_f <- quantile(ratio_coef_f, probs = c(0.05, 0.5, 0.95))
  print(CI_f)
}
conf_mm_m <- function(dat = gene1only, genename = "Egfr", fun = bootie){
  coef_out <- replicate(n = 100, fun(dat = dat, genename = genename))
  ratio_coef_m <- coef_out[11:13]/coef_out[15:17]
  CI_m <- quantile(ratio_coef_m, probs = c(0.05, 0.5, 0.95))
  print(CI_m)
}

```

# Vein Epistasis 


A genetic interaction or epistasis can be defined in its simplest form as relationships between genes. However, this term can mean different things depending on your field and questions. 
For this work, we will use this term to describe the interaction of genes at different loci that affect a phenotype.  Specifically, the interaction of a secondary mutation that enhances a focal mutation. 

## Scientific Question: 
##Does the magnitude of genetic interactions change in a background dependent manner?

### What does this mean? 

The phenotypic consequences of individual mutations are modulated by the wild-type genetic background in which they occur.The influence of wild-type genetic backgrounds also extends to interactions among mutations, but this needs to be explored further. 


### How are we approaching this problem? 

Examining epistatic interactions are experimentally, statistically, and computationally challenging and unfortunately there is no agreed upon way on how to handle such challenges. As such, our goal for this project is to use 4 different approaches to analyse our scientific question. This allows for us to explore different methods and see which we think is the best approach to answer our question. The best meaning taking into consideration overall fit of models, efficiency, output that allows for interpretation and so on. 


### Dataset 

* Data obtained from set of experimental crosses and data collected by Andrew Victory while an undergraduate at MSU in the Dworkin lab.
* Looking at known genetic interactions by observing vein lengths in double and triple mutants in two common strains of Drosophila, Sam and Ore. 
* All animals are also homozygous for a rhomboid mutation.

Aos = argos,                                    S = star,
Bs = blistered,                                 Spi = spitz,
Egfr = epidermal growth factor receptor

##Statistical Question:
## Is there variance in vein length that can not be explained by the effect of background and the effect of gene1 and/or gene2? 

###Choices we made: 

* All vein lengths are scaled by overall wing length. 
* We are using subsets of our data for most analysis - either individuals that are wild type at gene2 or individuals that are egfr gene 1 mutants and have a variant at gene 2

##Our plan

1. PCA to ask if the variance in vein lengths is correlated to inform the limitations in our other analysis
2. Supervein analysis - if changes in all of the vein lengths are correlated, can we just add them together to create a ‘supervein’
3. Mixed model approach - a mixed model with a repeated measure of each vein for each individual.
4. Multivariate mixed model - this approach is similar to approach 3. We want to compare results and ask which model is easier to interpret in the context of our biological question. 

##Cleandata 

```{r clean}
Vein_epistasis <- read_csv("~/Dropbox/KatiePelletier/statsProjectStuff/RhomboidEpistasisWingMeasures_Fixes_2019.csv")
#Change second gene to say wt if it does not have a second mutation -> get rid of NA in this column 
Vein_epistasis$gene2[is.na(Vein_epistasis$gene2)] <- "wt"
#change relevant columns to factors 
columns_to_factors <- c("background", "gene1", "gene2", "rhomboid", "sex", "individual")
Vein_epistasis[,columns_to_factors] <- lapply(Vein_epistasis[,columns_to_factors], as.factor)
#omit NAs - indicates dead flies 
Vein_epistasis <- na.omit(Vein_epistasis)

#checking for any obvious outliers
#hist
Vein_epistasis %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
  facet_wrap(~ key, scales = "free") +
  geom_histogram()

#might be outliers in L5 and WL 

#boxplot for L5
(ggplot(Vein_epistasis,aes(gene1, L5))
  + geom_boxplot()
  + stat_sum(colour="darkgray",alpha=0.5)
  + scale_size(breaks=1:2, range=c(3,6))
)

#remove outlier 
Vein_epistasis<- subset(Vein_epistasis, Vein_epistasis[11]<5)

#double check outlier has been removed 

(ggplot(Vein_epistasis,aes(gene1, L5))
  + geom_boxplot()
  + stat_sum(colour="darkgray",alpha=0.5)
  + scale_size(breaks=1:2, range=c(3,6))
)
#boxplot for WL

(ggplot(Vein_epistasis,aes(gene1, WL))
  + geom_boxplot()
  + stat_sum(colour="darkgray",alpha=0.5)
  + scale_size(breaks=1:2, range=c(3,6))
)

#remove outlier 

Vein_epistasis<- subset(Vein_epistasis, Vein_epistasis[12]>4)

#double check outlier has been removed 

(ggplot(Vein_epistasis,aes(gene1, WL))
  + geom_boxplot()
  + stat_sum(colour="darkgray",alpha=0.5)
  + scale_size(breaks=1:2, range=c(3,6))
)
#For some reason all the lenghts were created as factors and not numberic variables. 

Veins_numeric <- c("L2", "L3", "L4", "L5", "WL")
Vein_epistasis[,Veins_numeric] <- lapply(Vein_epistasis[,Veins_numeric], as.numeric)

#scale vein lengths by overall wing lengths 
Vein_epistasis_scale <- (Vein_epistasis
  %>% mutate(L2s = L2/WL)
  %>% mutate(L3s = L3/WL)
  %>% mutate(L4s = L4/WL)
  %>% mutate(L5s = L5/WL)
)

wings <- Vein_epistasis_scale

#We are choosing to simplify our analysis by only looking at interaction between rho and one gene (gene1 in dataset). Because of this, we want the gene2 position to always be wildtype
gene1only <- subset(wings, gene2 == "wt")

#We don't need to write out the data in this case, but will keep the code to do so 

#write.csv(Vein_epistasis, file ="Vein_epistasis_clean.csv")
#save(wings, 
#     file = "clean_epistasis.rda")
```

##PCA 

```{r PCA}
#I want to do a PC analysis on the vein data to ask the question:
#is it a good idea to consider each vein seperatley? Can we consider them all together as one 'supervein' 
#If there is not a lot of varriation within PCs, we can conclude that all four veins respond in the same way to a genetic pertubation
#Using the scaled values because allometric effects have been controled for. 
#Decided to leave the scaling set as false (defult) because I had already scaled all the vein lengths to the length of the wing.
#I did also run these tests with the resacling step in the PC decomposition and the same pattrens were observed.


gene1PCs <- prcomp(gene1only[,13:16], scale. = F)
gene1_combined <- data.frame(gene1only, gene1PCs$x[,1:4])

#Scree plot for the PCs is not particularly infomative because it is unscaled, looking at the summary is easier.
plot(gene1PCs)
summary(gene1PCs)

#We also want to look at the loadings for the PCs
#It looks like the genes are not corrolated with eachother, indicating that the supervein analysis is probably not the best choice. We also show this below when we try to do it. 
gene1PCs$rotation

#This is also clear when the PCs are plotted against one another. 
#This function plots the specified PCs with colour indicating the background and shape indicating sex

#There is a seperation based on background, but not on sex. 
PC_plot(x = gene1_combined$PC1, y = gene1_combined$PC2)
PC_plot(x = gene1_combined$PC2, y = gene1_combined$PC3)
PC_plot(x = gene1_combined$PC3, y = gene1_combined$PC4)

#I don't expect any PC to be particularly corrolated with size because we scaled the vein length relitive to overall wing length but I want to check

#Ore has a positive slope, but the data has a large spread so I'm not going to worry too much 
#This only looks like there is a pattren for PC1
size_cor(PC = gene1_combined$PC1)
size_cor(PC = gene1_combined$PC2)
size_cor(PC = gene1_combined$PC3)
size_cor(PC = gene1_combined$PC4)

##Now we want to use the PCs in place of vein length for the model. 
##Going to use PC1 as the response variable because It captures the greatest proportion of the varriance and shows the most clear distiction between backgrounds

#Staring with the model we will use below in our analysis.
#This is the maximal model
#lmer gave a signular fit. 
PC1_mod <-  glmmTMB(PC1 ~ background*gene1*sex + (1|rep), data = gene1_combined)


PC1_resid <- simulateResiduals(PC1_mod)
plot(PC1_resid)
#This looks bad.
#Looked at the residual varriance for the random effect in our model, it is small so that is probably not the reason our fit looks so bad. 
summary(PC1_mod)
#However, there was no model with a better fit that we could find. 
#This is proabably because the distribution of PC1 is skewed and does not fit our assumption of normality. 
hist(gene1_combined$PC1)

#For now, we are going to continue with this model, although we can not put a lot of stock in the outcome. 
#We could revisit this using a baysian framework in the future to model these effects. 

#In this case, there is no significant interaction with sex, this indicates that we may be able to take sex out of our analysis to simplify it. 
Anova(PC1_mod)

#Looking at the contrast between the backgrounds, based on the genes. 
pairwise_PC1 <- emmeans(PC1_mod, pairwise ~ background|gene1)

#plotting the marginal means, faceted by gene, with 95% confidence intrivals. 
#These are clearly diffrent, although this does not give a magnitude of effect 
plot(pairwise_PC1)

#Here we are able to see the magnitude of diffrence between the two means. 
#This still doesn't have a CI on the magnitude of diffrence but can use SE as a proxy of a measure of varriance. 

pairwise_PC1

##Going to reuse the bootstraping function we wrote to compare coefficents to compare the magnitude of change for PC1

#This is diffrent from one, therefore the magnitude of the episatic effect between Egfr and rho is diffrent in the Ore and Sam background for females. 
conf_bootie_f(dat = gene1_combined, genename = "Egfr", vein = gene1_combined$PC1, fun = bootie)

#This is not diffrent from one, therefore the magnitude of the episatic effect between Egfr and rho is not in the Ore and Sam background for males. These conf. intrivals are also huge
conf_bootie_m(dat = gene1_combined, genename = "Egfr", vein = gene1_combined$PC1, fun = bootie)

#This is not what we predicted, but it is an interesting biological result that can be followed up on. Why do we observe sexual dimporphism in background dependance of epistatic interactions?

#I will look at the other genes as well. 
#No change here 
conf_bootie_f(dat = gene1_combined, genename = "S", vein = gene1_combined$PC1, fun = bootie)
conf_bootie_m(dat = gene1_combined, genename = "S", vein = gene1_combined$PC1, fun = bootie)

#No change 
conf_bootie_f(dat = gene1_combined, genename = "Spi", vein = gene1_combined$PC1, fun = bootie)
conf_bootie_m(dat = gene1_combined, genename = "Spi", vein = gene1_combined$PC1, fun = bootie)

#No change
conf_bootie_f(dat = gene1_combined, genename = "Bs", vein = gene1_combined$PC1, fun = bootie)
conf_bootie_m(dat = gene1_combined, genename = "Bs", vein = gene1_combined$PC1, fun = bootie)

#No change
conf_bootie_f(dat = gene1_combined, genename = "Aos", vein = gene1_combined$PC1, fun = bootie)
conf_bootie_m(dat = gene1_combined, genename = "Aos", vein = gene1_combined$PC1, fun = bootie)

#Although this model isn't great and I wouldn't put too much stock in these results as they are right now. 
```

##Supervein 

```{r Supervein}
#make new column with total vein length
wings$supervein <- wings$L2s + wings$L3s + wings$L4s  + wings$L5s
#subset so that gene2 equals wild-type. Only want to look at double mutants for now. 
gene1only <- subset(wings, gene2 == "wt")
#Look at data
ggplot(data = gene1only, aes(x=gene1, y=supervein)) + geom_boxplot(aes(fill=background))
#look at data to see if linear looks like it should reasonable fit 
hist(wings$supervein)
qqnorm(wings$supervein)
#mixed linear model using glmmTMB package. Singular fit when using lmer. 

lmm1_sv <- glmmTMB(supervein ~ background*gene1*sex + (1|rep), data = gene1only)
#check model 

supervein_lmm1 <- simulateResiduals(lmm1_sv)
plot(supervein_lmm1)
#there is deviation for expected distrubution 
#simulation outliers are highlighted as red stars. data points that are outside the range of simulated values. 
#fit doesn't look aweome which is expected - indicates this is not a good way to handle data.
summary(lmm1_sv)
#genes have effect which is what we would expect. Interesting that sex does as overall wing size has been scaled for, and even more interesting that interaction between background and sex has effect (meaning effect of sex behaves differently in background). But we need to compare magnitudes now. 
 
#inference
Anova(lmm1_sv)
#strong support there is sign effect of background and gene1 on supervein (This is predicted). What we care about is there looks to be significant effect of interaction of gene1 and background and background and sex. 

#visual of effects

(eff <- allEffects(lmm1_sv))
plot(eff)
 
```

Supervein approach is not a good approach. This was anticipated but preformed as an exploratory exercise. 

##What does model liner mixed model look like for each vein individually? This is a substitute for supervein approach. 

###Vein 2
```{r vein 2}

lmm1_L2s <- glmmTMB(L2s ~ background*gene1*sex + (1|rep), data = gene1only)
summary(lmm1_L2s)
Anova(lmm1_L2s)

emmip(lmm1_L2s, sex ~ background|gene1, CIs = T)


#A comparison of the two coeficents between backgrounds. Females and males are considered seperatley 
#Functions can be found at the top of this document. Bootstrap analysis with 100 replicates to get CI. 
#A value of 1 would mean there is no change in magnitude between the two backgrounds. 

#Change for females but not males for the interaction between Egfr and rho 
conf_bootie_f(dat = gene1_only, genename = "Egfr", vein = gene1_combined$L2s, fun = bootie)
conf_bootie_m(dat = gene1_only, genename = "Egfr", vein = gene1_combined$L2s, fun = bootie)

#Again, a change for females but not for males for S. 
conf_bootie_f(dat = gene1_only, genename = "S", vein = gene1_combined$L2s, fun = bootie)
conf_bootie_m(dat = gene1_only, genename = "S", vein = gene1_combined$L2s, fun = bootie)

#No change for Spi in either sex 
conf_bootie_f(dat = gene1_only, genename = "Spi", vein = gene1_combined$L2s, fun = bootie)
conf_bootie_m(dat = gene1_only, genename = "Spi", vein = gene1_combined$L2s, fun = bootie)

#No change for Bs
conf_bootie_f(dat = gene1_only, genename = "Bs", vein = gene1_combined$L2s, fun = bootie)
conf_bootie_m(dat = gene1_only, genename = "Bs", vein = gene1_combined$L2s, fun = bootie)

#No change for Aos
conf_bootie_f(dat = gene1_only, genename = "Aos", vein = gene1_combined$L2s, fun = bootie)
conf_bootie_m(dat = gene1_only, genename = "Aos", vein = gene1_combined$L2s, fun = bootie)
```

###vein 3
```{r vein 3}
lmm1_L3s <- glmmTMB(L3s ~ background*gene1*sex + (1|rep), data = gene1only)
summary(lmm1_L3s)
Anova(lmm1_L3s)
emmip(lmm1_L3s, sex ~ background|gene1, CIs = T)

#We could do the same tests as above to compare effect sizes, but again, this would be a lot of the same code over and over again. Insted we will just compare for egfr for the rest of this section. 

#No change in either sex, not what we saw for L2. 
conf_bootie_f(dat = gene1_only, genename = "Egfr", vein = gene1_combined$L3s, fun = bootie)
conf_bootie_m(dat = gene1_only, genename = "Egfr", vein = gene1_combined$L3s, fun = bootie)
```

###vein 4
```{r vein 4}
lmm1_L4s <- glmmTMB(L4s ~ background*gene1*sex + (1|rep), data = gene1only)
summary(lmm1_L4s)
Anova(lmm1_L4s)
emmip(lmm1_L4s, sex ~ background|gene1, CIs = T)

#Again, no change. 
conf_bootie_f(dat = gene1_only, genename = "Egfr", vein = gene1_combined$L4s, fun = bootie)
conf_bootie_m(dat = gene1_only, genename = "Egfr", vein = gene1_combined$L4s, fun = bootie)
```

###vein 5
```{r vein 5}
lmm1_L5s <- glmmTMB(L5s ~ background*gene1*sex + (1|rep), data = gene1only)
summary(lmm1_L5s)
Anova(lmm1_L5s)
emmip(lmm1_L5s, sex ~ background|gene1, CIs = T)

#No change again. 
conf_bootie_f(dat = gene1_only, genename = "Egfr", vein = gene1_combined$L5s, fun = bootie)
conf_bootie_m(dat = gene1_only, genename = "Egfr", vein = gene1_combined$L5s, fun = bootie)
```


##Multiple measurements approach 
```{r mixed model}
##Going to try to fix a model using each vein length as a repeated measure on each individual

##first, I need to modify the data so that each individual has 4 vein measurements. 
gene1only_long <- (gene1only
          %>% select(-c("L2","L3","L4","L5", "WL", "supervein")) #removes the unsacled values so I can't mess up and use these also removing the supervein. 
          %>% mutate(units=factor(1:n())) #each individual becomes a unit 
          %>% gather(vein, scalelength, -c(units,rep, sex,background, gene1, gene2, individual, rhomboid))
)

gene1only_long$vein <- as.factor(gene1only_long$vein)

#Fitting the model (same as one used above)
mod1_S <- lmer(scalelength ~ vein:(background*gene1*sex)-1 + 
                 (1|rep) + (vein - 1|units), 
               data = gene1only_long, control = lmerControl(optCtrl=list(ftol_abs=1e-10),
                                                            optimizer="bobyqa",
                                                            check.nobs.vs.nlev="ignore",
                                                            check.nobs.vs.nRE="ignore"))

#No sigular fit, good! 
all(abs(getME(mod1_S,"theta"))>1e-4)

#Looking for corolations in the data 
#This is more evidence for why supervein was a bad idea, there is a moderate assosoiation between L3 and L4 but no other veins. 
corplot <- VarCorr(mod1_S)
diag(corplot$units) <- diag(corplot$units)+sigma(mod1_S)^2
corrplot.mixed(cov2cor(corplot$units),upper="ellipse")

#creating a coefficent plot 
cc <- tidy(mod1_S,effect="fixed") %>%
  separate(term,into=c("vein", "fixeff"),extra="merge",
                  remove=FALSE)


#This is really hard to look at 
##I think the most important take awawy here is that the 'epistatic' term is diffrent for each vein. 
#This was shown more clearly with the corr plot. 
dwplot(cc)+facet_wrap(~fixeff,scale="free",ncol=4)+
  geom_vline(xintercept=0,lty=2)

#Here all the genes seem to follow the same pattren of effects across the sexes and the backgrounds. 
eff_lmm <- allEffects(mod1_S)
plot(eff_lmm)

#Inferance
#This tells us that the interaction between vein, background, gene and sex is significant. Although, this tells us nothing about the magnitude of change between the backgrounds. 
Anova(mod1_S)

#To compare sizes of effects, we would have to do a bootstrap analyis similar to what was done above, however, this would take longer and be more complicated because of the number of factors included in this model.
#This takes the average of L2, L3, and L4 coefficents. L5 is removed from this analysis because something weird is going on with it. 

#Change for males but not females. 
conf_mm_f(dat = gene1only_long, genename = "Egfr", fun = bootie_mm)
conf_mm_m(dat = gene1only_long, genename = "Egfr", fun = bootie_mm)

#Change for males but not females. 
conf_mm_f(dat = gene1only_long, genename = "S", fun = bootie_mm)
conf_mm_m(dat = gene1only_long, genename = "S", fun = bootie_mm)

#No change. 
conf_mm_f(dat = gene1only_long, genename = "Spi", fun = bootie_mm)
conf_mm_m(dat = gene1only_long, genename = "Spi", fun = bootie_mm)

#No change.
conf_mm_f(dat = gene1only_long, genename = "Bs", fun = bootie_mm)
conf_mm_m(dat = gene1only_long, genename = "Bs", fun = bootie_mm)

#No change.
conf_mm_f(dat = gene1only_long, genename = "Aos", fun = bootie_mm)
conf_mm_m(dat = gene1only_long, genename = "Aos", fun = bootie_mm)
```


##Multivariate 

```{r multivariate}
#variance covariance matrix 

cov(gene1only[ ,13:16])

#visual variance covariance matrix
pairs(gene1only[, 13:16],
      pch = ".", gap = 0)


scatterplotMatrix( ~ L2s + L3s + L4s + L5s | interaction(gene1, background), 
                   ellipse = TRUE, data = gene1only, gap = 0,
                   plot.points = T, pch = 20, cex  = 0.5)

#examine the eigenvalues of the covariance matri - you don't want really small values
#check to see if the covariance matrix was not of full rank

eig_vals <- svd(cov(gene1only[, 13:16]))$d
eig_vals
det(cov(gene1only[, 13:16]))

mlm_fit1 <- lm(as.matrix(gene1only[,13:16]) ~ background*gene1*sex, data = gene1only)
class(mlm_fit1)

summary(manova(mlm_fit1))

summary(manova(mlm_fit1), 
        test = "Wilks")

Anova(mlm_fit1)

coef1 <- coef(mlm_fit1)

bootiefunction_mlm <-  function(dat=gene1only, genename = "Egfr") {
  new_dat <- gene1only[sample(nrow(gene1only), nrow(gene1only), replace = T),  ]
  mod_new <- lm(as.matrix(new_dat[,13:16]) ~ background*sex, 
                data = new_dat,
                subset = gene1only$gene1 == genename)
  return(coef(mod_new))
}

coef_out <- replicate(n = 100, bootie())
dim(coef_out)
#l2
ratio_coef_f_l2 <- coef_out[2,1,]/coef_out[1,1,]
quantile(ratio_coef_f, probs = c(0.5, 0.5, 0.95))
ratio_coef_m_L2 <- coef_out[4,1,]/coef_out[3,1,]
quantile(ratio_coef_m, probs = c(0.05, 0.5, 0.95))
#L3
ratio_coef_f_l3 <- coef_out[2,2,]/coef_out[1,2,]
quantile(ratio_coef_f, probs = c(0.5, 0.5, 0.95))
ratio_coef_m_L3 <- coef_out[4,2,]/coef_out[3,2,]
quantile(ratio_coef_m, probs = c(0.05, 0.5, 0.95))
#L4
ratio_coef_f_l3 <- coef_out[2,3,]/coef_out[1,3,]
quantile(ratio_coef_f, probs = c(0.5, 0.5, 0.95))
ratio_coef_m_L3 <- coef_out[4,3,]/coef_out[3,3,]
quantile(ratio_coef_m, probs = c(0.05, 0.5, 0.95))
#L5
ratio_coef_f_l3 <- coef_out[2,5,]/coef_out[1,5,]
quantile(ratio_coef_f, probs = c(0.5, 0.5, 0.95))
ratio_coef_m_L3 <- coef_out[4,5,]/coef_out[3,5,]
quantile(ratio_coef_m, probs = c(0.05, 0.5, 0.95))

```


For this assignment, we took four diffrent approaches to answer the same problem to investigate the efficency and ease of understanding our results. 
Using PC analysis, we were able to demonstrate a diffrence between the two backgrounds by clustering of the PCs. However, our model for this analysis was not perfect and would have to be improved upon before any firm conclusions can be made. That being said, this was an efficent way to gather the varriance captured but all 4 measured veins and only need to do a single analysis on PC1 to explain most of the caputred varriation. 
We also tried to add all veins together to create a 'supervein', using the assumption that changes to vein length of all four veins would be corrolatated. This was not sucessful because our assumption was not correct. Insted, we chose to fit a model to each vein seperatley. This was able to produce an answer that we could then inturput in the context of our biological question. However, this approach was quite cumbersome. 
We then fit a model using each vein as a measurement for an individual. This was simlilar to the above approach but allowed us to fit the model for all the veins at the same time as we do not care about changes to individual veins, just the overall effect. This model also produced a result that was easy to understand with less code than the above model. Although, this gave conflicting results to the other two tests. I'm not sure if this is a problem in the code or a diffrent approach giving a diffrent answer. 
