---
title: "Final Project"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Vein Epistasis 


A genetic interaction or epistasis can be defined in its simplest form as relationships between genes. However, this term can mean different things depending on your field and questions. 
For this work, we will use this term to describe the interaction of genes at different loci that affect a phenotype.  Specifically, the interaction of a secondary mutation that enhances a focal mutation. 

## Scientific Question: 
##Does the magnitude of genetic interactions change in a background dependent manner?

### What does this mean? 

The phenotypic consequences of individual mutations are modulated by the wild-type genetic background in which they occur.The influence of wild-type genetic backgrounds also extends to interactions among mutations, but this needs to be explored further. 


### How are we approaching this problem? 

Examining epistatic interactions are experimentally, statistically, and computationally challenging and unfortunately there is no agreed upon way on how to handle such challenges. As such, our goal for this project is to use 4 different approaches to analyse our scientific question. This allows for us to explore different methods and see which we think is the best approach to answer our question. The best meaning taking into consideration overall fit of models, efficiency, output that allows for interpretation and so on. 


### Dataset 

* Data obtained from set of experimental crosses and data collected by Andrew Victory while an undergraduate at MSU in the Dworkin lab.
* Looking at known genetic interactions by observing vein lengths in double and triple mutants in two common strains of Drosophila, Sam and Ore. 
* All animals are also homozygous for a rhomboid mutation.

Aos = argos,                                    S = star,
Bs = blistered,                                 Spi = spitz,
Egfr = epidermal growth factor receptor

##Statistical Question:
## Is there variance in vein length that can not be explained by the effect of background and the effect of gene1 and/or gene2? 

###Choices we made: 

* All vein lengths are scaled by overall wing length. 
* We are using subsets of our data for most analysis - either individuals that are wild type at gene2 or individuals that are egfr gene 1 mutants and have a variant at gene 2

##Our plan

1. PCA to ask if the variance in vein lengths is correlated to inform the limitations in our other analysis
2. Supervein analysis - if changes in all of the vein lengths are correlated, can we just add them together to create a ‘supervein’
3. Mixed model approach - a mixed model with a repeated measure of each vein for each individual.
4. Multivariate mixed model - this approach is similar to approach 3. We want to compare results and ask which model is easier to interpret in the context of our biological question. 

##Cleandata 

```{r clean}
library("tidyverse")
Vein_epistasis <- read_csv("GeneticBackgroundVeinMutantInteractionsFinal.csv")
#Change second gene to say wt if it does not have a second mutation -> get rid of NA in this column 
Vein_epistasis$gene2[is.na(Vein_epistasis$gene2)] <- "wt"
#change relevant columns to factors 
columns_to_factors <- c("background", "gene1", "gene2", "rhomboid", "sex", "individual")
Vein_epistasis[,columns_to_factors] <- lapply(Vein_epistasis[,columns_to_factors], as.factor)
#omit NAs - indicates dead flies 
Vein_epistasis <- na.omit(Vein_epistasis)

#checking for any obvious outliers
#hist
Vein_epistasis %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
  facet_wrap(~ key, scales = "free") +
  geom_histogram()

#might be outliers in L5 and WL 

#boxplot for L5
(ggplot(Vein_epistasis,aes(gene1, L5))
  + geom_boxplot()
  + stat_sum(colour="darkgray",alpha=0.5)
  + scale_size(breaks=1:2, range=c(3,6))
)

#remove outlier 
Vein_epistasis<- subset(Vein_epistasis, Vein_epistasis[11]<5)

#double check outlier has been removed 

(ggplot(Vein_epistasis,aes(gene1, L5))
  + geom_boxplot()
  + stat_sum(colour="darkgray",alpha=0.5)
  + scale_size(breaks=1:2, range=c(3,6))
)
#boxplot for WL

(ggplot(Vein_epistasis,aes(gene1, WL))
  + geom_boxplot()
  + stat_sum(colour="darkgray",alpha=0.5)
  + scale_size(breaks=1:2, range=c(3,6))
)

#remove outlier 

Vein_epistasis<- subset(Vein_epistasis, Vein_epistasis[12]>4)

#double check outlier has been removed 

(ggplot(Vein_epistasis,aes(gene1, WL))
  + geom_boxplot()
  + stat_sum(colour="darkgray",alpha=0.5)
  + scale_size(breaks=1:2, range=c(3,6))
)
#For some reason all the lenghts were created as factors and not numberic variables. 

Veins_numeric <- c("L2", "L3", "L4", "L5", "WL")
Vein_epistasis[,Veins_numeric] <- lapply(Vein_epistasis[,Veins_numeric], as.numeric)

#scale vein lengths by overall wing lengths 
Vein_epistasis_scale <- (Vein_epistasis
  %>% mutate(L2s = L2/WL)
  %>% mutate(L3s = L3/WL)
  %>% mutate(L4s = L4/WL)
  %>% mutate(L5s = L5/WL)
)

wings <- Vein_epistasis_scale
write.csv(Vein_epistasis, file ="Vein_epistasis_clean.csv")
save(wings, 
     file = "clean_epistasis.rda")
```

##PCA 

```{r PCA}


```

##Supervein 

```{r Supervein}
#make new column with total vein length
wings$supervein <- wings$L2s + wings$L3s + wings$L4s  + wings$L5s
#subset so that gene2 equals wild-type. Only want to look at double mutants for now. 
gene1only <- subset(wings, gene2 == "wt")
#Look at data
library(ggplot2)
ggplot(data = gene1only, aes(x=gene1, y=supervein)) + geom_boxplot(aes(fill=background))
library(lme4)
#look at data to see if linear looks like it should reasonable fit 
hist(wings$supervein)
qqnorm(wings$supervein)
#mixed linear model using glmmTMB package. Singular fit when using lmer. 
library(glmmTMB)
lmm1_sv <- glmmTMB(supervein ~ background*gene1*sex + (1|rep), data = gene1only)
summary(lmm1_sv)
#genes have effect which is what we would expect. Interesting that sex does as overall wing size has been scaled for, and even more interesting that interaction between background and sex has effect (meaning effect of sex behaves differently in background). 
#check model 
library(DHARMa)
supervein_lmm1 <- simulateResiduals(lmm1_sv)
plot(supervein_lmm1)
#there is deviation for expected distrubution 
#simulation outliers are highlighted as red stars. data points that are outside the range of simulated values. 
#fit doesn't look aweome which is expected - indicates this is not a good way to handle data. 
#inference
library(car)
Anova(lmm1_sv)
#strong support there is sign effect of background and gene1 on supervein (This is predicted). What we care about is there looks to be significant effect of interaction of gene1 and background and background and sex. 

#visual of effects
library(effects)
(eff <- allEffects(lmm1_sv))
plot(eff)
 
```

Supervein approach is not a good approach. This was anticipated but preformed as an exploratory exercise. 

##What does model liner mixed model look like for each vein individually? This is a substitute for supervein approach. 

###Vein 2
```{r vein 2}
gene1only$gene1 <- relevel(gene1only$gene1, "wt")
library(lme4)
lmm1_L2s <- glmmTMB(L2s ~ background*gene1*sex + (1|rep), data = gene1only)
summary(lmm1_L2s)
Anova(lmm1_L2s)
library(emmeans)
emmip(lmm1_L2s, sex ~ background|gene1, CIs = T)
```

###vein 3
```{r vein 3}
lmm1_L3s <- glmmTMB(L3s ~ background*gene1*sex + (1|rep), data = gene1only)
summary(lmm1_L3s)
Anova(lmm1_L3s)
emmip(lmm1_L3s, sex ~ background|gene1, CIs = T)

```

###vein 4
```{r vein 4}
lmm1_L4s <- glmmTMB(L4s ~ background*gene1*sex + (1|rep), data = gene1only)
summary(lmm1_L4s)
Anova(lmm1_L4s)
emmip(lmm1_L4s, sex ~ background|gene1, CIs = T)
```

###vein 5
```{r vein 5}
lmm1_L5s <- glmmTMB(L5s ~ background*gene1*sex + (1|rep), data = gene1only)
summary(lmm1_L5s)
Anova(lmm1_L5s)
emmip(lmm1_L5s, sex ~ background|gene1, CIs = T)

```

```{r bootie}

```


##Mixed Model Approach 
```{r mixed model}


```


##Multivariate 

```{r multivariate}
load("clean_epistasis.rda")

gene1only <- subset(wings, gene2 == "wt")

#variance covariance matrix 

cov(gene1only[ ,13:16])

#visual variance covariance matrix
pairs(gene1only[, 13:16],
      pch = ".", gap = 0)


library(car)
scatterplotMatrix( ~ L2s + L3s + L4s + L5s | interaction(gene1, background), 
                   ellipse = TRUE, data = gene1only, gap = 0,
                   plot.points = T, pch = 20, cex  = 0.5)

#examine the eigenvalues of the covariance matri - you don't want really small values
#check to see if the covariance matrix was not of full rank

eig_vals <- svd(cov(gene1only[, 13:16]))$d
eig_vals
det(cov(gene1only[, 13:16]))

mlm_fit1 <- lm(as.matrix(gene1only[,13:16]) ~ background*gene1*sex, data = gene1only)
class(mlm_fit1)

summary(manova(mlm_fit1))

summary(manova(mlm_fit1), 
        test = "Wilks")

Anova(mlm_fit1)

coef1 <- coef(mlm_fit1)


mlm_coef_f <- coef1[9,1]/ coef1[3,1]
mlm_coef_f

mlm_coef_m <- coef1[20,1]/ coef1[15,1]
mlm_coef_m

